---
title: "Exercise 1"
author: "Clark, Zach, Zargham"
date: "2/5/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(FNN)
library(mosaic)
library(ggplot2)
```

## Patterns of Delay & Cancellation at Austin Bergstorm Airport

We look at the data collected on flights at ABIA to check for intresting patterns. Some of those patterns include the delay or cancellation of flights with respect to different carriers, days of the month and days of the week. 

The data includes flights into and out of ABIA. In this exercise we focus on the patterns of departure delay for flights going out of ABIA. For this we create a new subset that includes only those flights which depart from ABIA. Also we consider any flights that depart minutes before their scheduled time to be on-time and convert negative values to zeros to give more realistic picture and without affecting average values for departure delay. 
```{r}
abia = read.csv('ABIA.csv')
naustin = subset(abia, Origin=="AUS")
naustin$DepDelay = ifelse(naustin$DepDelay < 0, 0, naustin$DepDelay)
```

## Carrier Wise Patterns of Delays
The first pattern we try to explore is the average departure delay for different carriers and see which carriers have more delays on average. This will show the performance of carriers. 

```{r}
carrier_sum = naustin %>%
      group_by(UniqueCarrier)  %>%  # group the data points by model nae
   summarize(DD.mean=mean(DepDelay ,na.rm=TRUE)) 

ggplot(carrier_sum, aes(x=reorder(UniqueCarrier,DD.mean), y=DD.mean)) + 
   geom_bar(stat='identity') + 
   coord_flip()+
   labs(title="Average Departure Delay per Carrier", 
        caption="Source: ABIA",
        x="Carrier",
        y = "Delay in minutes")
```



When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
