---
title: "Question 1"
output: 
  md_document:
        variant: markdown_github
---

```{r include=FALSE, cache= TRUE}

library(mosaic)
library(tidyverse)
library(Metrics)
library(gamlr)
library(margins)
library(rpart)
library(caret)
library(knitr)
library(randomForest)
library(gbm)
greenbuildings = read.csv("greenbuildings.csv", header=TRUE)
greenbuildings = na.omit(greenbuildings)
gb = greenbuildings[which(greenbuildings$leasing_rate != 0),]
gb$size = gb$size/1000

```

# Assignment 3
# Group Members: Clark Granger, Zachary Carlson and Zargham Khan

## Question 1. Predictive model for Green Buildings

Using the data on 7,984 commercial rental properties from the United States, the goal is to build a predictive model for rental income per square foot. To build the model we have several key features of each property such as size in square foot, leasing rate, number of stories, amenities, among others. We also have information about building's "neiborhood", weather in the area and kind of contract that the property offers. Beyond the amount of variables we have, we are specially interested in quantifying the average change in rental income per square foot  associated with green certification, holding other features of the building constant. In this dataset, 685 properties have been awarded either LEED or EnergyStar certification as a green building.

To adress this problem we use different approachs we have seen in class. The first approach is from the point of view of linear models, including model regularization techniques and stepwise selection. In the second approach we use a Lasso regression model and cross validation, to try to find a best model. Finally we implement tree models, random forest and boosting. 

### Approach #1 Linear models and stepwise selection 

To start we cleaned our database eliminating the properties that are without occupation and the entries with missing data. After that, we proceed creating a model that include variables we consider are relevant to explain the rent price. It is important to remark that we decided working with the variable that summarize green rating which takes value of 1 if the property has at least one green certifification and zero otherwise. Then, the first model we fitted is a linear one, called manual, that include most of the variables available, we mainly dropped variables that seemed to do not explain to much in our model. These variableswere property ID, precipitation days, cold and hot degree days, however we included total degrees days. After the first model, we regress a second one with the same varianles but now considering all the interactions between covariates. Having our two first models, the next step was working with the function step to look for a model that improves the AIC of the previous ones. To do that we created a null model and ran the function step to find a better model based on the scope of variables of the two originals models. After we got the new 2 models, we splitted up our sample in a train and test subsamples. To validate our models we generated out of sample RMSEs boostrapping our data. 

```{r include=FALSE, cache= TRUE}

#############################################################
#####APPROACH 1 fitting Linear models and step selection ############
#############################################################

#### Null model####
m0 = lm(Rent~1, data=gb)
summary(m0)

#### Manual built model####
m1  = lm(Rent ~ . -CS_PropertyID-LEED-Energystar
          -Precipitation-cd_total_07-hd_total07, data=gb)
summary(m1)

#### Manual model allowing interactions ####
m2  = lm(Rent ~ (. -CS_PropertyID-LEED-Energystar
                 -Precipitation-cd_total_07-hd_total07)^2, data=gb)
summary(m2)

#### Step selection ####
m3 = step(m0, scope=formula(m1), dir="forward")
summary(m3)
formula(m3)

m4 = step(m0, scope=formula(m2), dir="forward")
summary(m4)
formula(m4)
marm4=margins(m4)

# Compare out of sample performance
n = nrow(gb)
n_train = round(0.8*n)  # round to nearest integer
n_test = n - n_train
rmse_lm = do(50)*{
    # re-split into train and test cases with the same sample sizes
  train_cases = sample.int(n, n_train, replace=FALSE)
  test_cases = setdiff(1:n, train_cases)
  gb_train = gb[train_cases,]
  gb_test = gb[test_cases,]
  
  # Fit to the training data
  # use `update` to refit the same model with a different set of data
  lm1 = update(m1, data=gb_train)
  lm2 = update(m2, data=gb_train)
  lm3 = update(m3, data=gb_train)
  lm4 = update(m4, data=gb_train)
  
  # Predictions out of sample
  yhat_test1 = predict(lm1, gb_test)
  yhat_test2 = predict(lm2, gb_test)
  yhat_test3 = predict(lm3, gb_test)
  yhat_test4 = predict(lm4, gb_test)
  
  c(rmse(gb_test$Rent, yhat_test1),
    rmse(gb_test$Rent, yhat_test2),
    rmse(gb_test$Rent, yhat_test3),
    rmse(gb_test$Rent, yhat_test4))
}

rmse.lm=colMeans(rmse_lm)
aic.lm=AIC(m1,m2,m3,m4)
bic.lm=BIC(m1,m2,m3,m4)
```

**Table 1. RSME, AIC and BIC for linear models**
```{r echo=FALSE, warning= FALSE, cache= TRUE}
AIC=aic.lm$AIC
BIC=bic.lm$BIC
DF=bic.lm$df

a=cbind(rmse.lm,AIC,BIC,DF)
colnames(a)<-c("RMSE","AIC","BIC","DF")
row.names(a) <- c("Model 1: Manual built","Model 2: Interactions
                        ","Model 3: Manual built + step","Model 4: Interactions + step")
table1=kable(a, digits=2, padding = 2)
table1
```

Table 1 summarizes the information of the four linear models we built. we can see that the model with the lowest RMSE is the model 4 generated by the step function based on the all interactions model. We can also see that this model has the second lowest AIC, only improved by model 2. However, we decided to include BIC criteria too, since it penalizes the number of coefficientes included in a model. With this criteria, the model 4 is the better, considering it is a more parsimonious model than model 2. 

### Approach #2  Lasso Regression: Variable Selection and Regularization  
```{r include=FALSE, cache= TRUE}
##############################################################
####### APPROACH 2 Lasso Regression and Regularization #######
##############################################################
rmse_lasso = do(10)*{
  # re-split into train and test cases with the same sample sizes
  train_cases = sample.int(n, n_train, replace=FALSE)
  test_cases = setdiff(1:n, train_cases)
  gb_train = gb[train_cases,]
  gb_test = gb[test_cases,]
  x_train= sparse.model.matrix(Rent ~ (.-CS_PropertyID-LEED-Energystar)^2, data=gb_train)[,-1] # do -1 to drop intercept!
  x_test= sparse.model.matrix(Rent ~ (.-CS_PropertyID-LEED-Energystar)^2, data=gb_test)[,-1]
  y=gb_train$Rent

###### CV Lasso best model ######
cvm5= cv.gamlr(x_train, y, nfold=10,  verb=TRUE)

# plot the out-of-sample deviance as a function of log lambda
# Q: what are the bars associated with each dot? 
plot(cvm5, bty="n")

## CV min deviance selection
scb.min = coef(cvm5, select="min")
lambda=log(cvm5$lambda.min)
coef.lasso=sum(scb.min!=0) 

yhat_test5 = as.numeric(predict(cvm5, x_test, select="min"))

c(rmse(gb_test$Rent,yhat_test5),lambda,coef.lasso)

}

info.lasso=colMeans(rmse_lasso)
lambda1=info.lasso[2]
coef.lasso1=info.lasso[3]

```  

The second approach is from the point of view of lasso regression. Working with the same data as the linear models, the only extraa step we needed to take before start with lasso models was the creation of sparse matrices for training and testing subsamples. The sparse matrices included the same variables of our manual linear model plus all the interactions, the last in order to have and equivalent scope of linear models. We went straight to use the cross validation function to find the best lasso model to predict rent. The cross validation function finds the lasso model tha minimized the lambda and drops all the coefficients that are close to be zero. We know when lambda is close to zero the betas found by lasso are essentially the least squares estimates. Thus, we implemented a cross validation lasso regression considering 10 folds and bostrapped again the process several times. 

**Table 2. Evaluation of Lasso Model**
```{r echo=FALSE, warning= FALSE, cache= TRUE}
info.lasso=as.data.frame(info.lasso)
colnames(info.lasso)<-c("Model 5: Lasso Model")
row.names(info.lasso) <- c("RMSE","Log Lambda",
                        "Number of coefficients")
table2=kable(info.lasso,digits=2)
table2
```

Table 2 contains the average results of the lasso cross validation. The model found has in average a log lambda of `r lambda1` and in average includes `r coef.lasso1` coefficients. However, the findings suggests that the lasso selected model does not improve the best linear model built in the previous section according with the boostrapped out of sample RSME computed. In graph 1, we show the evolution of a lasso regression for one case. We can appreciate how lasso approach found the minimum RMSE in a number of coefficients similar to the coefficients included in the best linear regression. The last reinforces that the linear model could be a good predictor of rent price.  

**Graph 1. Lasso model: Errror vs Log Lambda**
```{r echo=FALSE, warning= FALSE, cache= TRUE}
plot(cvm5, bty="n")
```


### Approach #3 Tree Models, Random Forest and Boosting 
```{r include=FALSE, cache= TRUE}

############################  
###### 3.A. Tree model #####
rmse_tree = do(50)*{
# re-split into train and test cases with the same sample sizes
train_cases = sample.int(n, n_train, replace=FALSE)
test_cases = setdiff(1:n, train_cases)
gb_train = gb[train_cases,]
gb_test = gb[test_cases,]

m6= rpart(Rent ~ .-CS_PropertyID-LEED-Energystar, method="anova",data=gb_train,
              control=rpart.control(minsplit=5, cp=1e-6, xval=10))
nm6 = length(unique(m6$where))

# calculate the cv error + 1 standard error
# the minimum serves as our threshold
err1se = m6$cptable[,'xerror'] + m6$cptable[,'xstd']
errth = min(err1se)

# now find the largest simplest tree that beats this threshold
m6$cptable[,'xerror'] - errth
stree=min(which(m6$cptable[,'xerror'] - errth < 0)) 
bestm6 = m6$cptable[stree,'CP']

###### Cross validation tree model ####
cvm6 = prune(m6, cp=bestm6)
length.btree=length(unique(cvm6$where))
plot(cvm6)
text(cvm6)

yhat_test6 = predict(cvm6, gb_test)
c(rmse(gb_test$Rent,yhat_test6),length.btree)
}
info.trees=colMeans(rmse_tree)
####################################
##### 3.B. Random Forest Model #####
m7 = randomForest(Rent ~ .-CS_PropertyID-LEED-Energystar,
                  data = gb_train, mtry = 5, ntree=500)
plot(m7)
yhat_test7 = predict(m7, gb_test)
rmse.rf=rmse(gb_test$Rent,yhat_test7)
varImp(m7) 
varImpPlot(m7,type=2,main="Variable Importance")

####################################
##### 3.C. Boosting Model #########

m8 = gbm(Rent ~ .-CS_PropertyID-LEED-Energystar,
         data = gb_train,n.trees=5000, shrinkage=.05)
yhat_test8= predict(m8, gb_test, n.trees=5000)
summary(m8)
rmse.boost=rmse(gb_test$Rent,yhat_test8)

```  

As a final approach, we work with tree models, random forest and boosting. First we worked with a simple tree model to predict rent price. About trees we known that they are good to manipulate large dataset and to ignore redundant data, we know they are not the best model to prediction though. A full tree was grown on the training set, with splitting continuing until a minimum bucket size of 5 was reached. This tree was pruned, and the tree size was chosen by 10-fold cross-validation. we repeated this approach several times by data boostrapping to generate out of sample RMSEs. 
After the simple tree, we proceeded with a random forest. This method fit many large trees to bootstrap-resampled versions of the training data
by relevance. We created 500 trees by random forest including a minimum of 5 features per bucket, and later we compute the out of sample RMSE. Finally, we performed a boosting model, that mainly fits several trees to reweighted versions of the training dataset and then classifies by weighted majority relevance. We fitted 5000 trees by this method with a shrinkage factor of 0.05. 

**Table 3. Tree, Random Forest and Boosting comparative**
```{r echo=FALSE, warning= FALSE, cache= TRUE}
table3=as.data.frame(c(info.trees[1],rmse.rf,rmse.boost))
colnames(table3)<-c("RMSE")
row.names(table3) <- c("Model 6: Tree model","Model 7:Random Forest",
                           "Model 8: Boosting")
table3=kable(table3,digits=2) 
table3
```

Table 3 summarizes the RMSE of the last tree model fitted. We can see that when the simple tree model does not improve the previous models, the random forest model and the boosting model certainly outperfom all the models. The best model we have built to predict the rent price per square foot so far is the random forest model with the least out of sample RMSE. However, now we face the problem that random forest model are certainly not interpretable and we want to know the partial effect of being a green building over price. Despite of this limitation random forest models allow us to get a importance meaure of each variable in the model.It just adds up how much the error decreases every time a variable is used in a split. This information is contained in graph 2. we can observe that the most importan variable is the rent price in the cluster where our property is located, followed by size and so on. 

**Graph 2. Varibale importance (Random Forest)**
```{r echo=FALSE, warning= FALSE, cache= TRUE}
varImpPlot(m7,type=2,main="")
```

```{r include=FALSE, cache= TRUE}
marm4
```
### Concluding Remarks

With data on 7,984 commercial rental properties from the United States, we built eight models to predict the rent price per square foot. The approaches implemented varie from linear models, going throug step selection, lasso regression until to tree models, random forest and boosting. The out of sample results suggets that the model with the lowest error is the random forest model with a RMSE of `r rmse.rf`. We also were requested to find the the average change in rental income per square foot  associated with green certification. Given the barrier we have to get this kind of interpretation from a random forest model. We computed the average marginal effect of being a green building from the best linear model we fitted (Model 4). The results suggest that having a green certification leads an increasing in 1.307 dollars per square foot in rent price. This result makes sense since green building properties have positive features such as lower energy compsumption levels, variable that according with our random forest model is important to explain rent price. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
library(ggplot2)
library(tidyverse)
library(dplyr)
library(foreach)
library(mosaic)
library(LICORS)
library(pander)
library(reshape2)
library(factoextra)

wine= read.csv("wine.csv")

```

## Question. 2: What causes what?

### Part A. 

Running a simple regression of crime on police will give us correlation between these two variables but it will not help us identify if one causes the other. It might be a case that crime affects police number because police numbers are apportioned by authority based on predetermined crime rates of areas. So a simple regression will be spurious and will not help us identify if higher crime leads to higher police number or vice versa.   

Even if the casual effect of police is negative but the regression might pick postive correlation due to high police in high crime areas which might lead people to infer that higher police causes higher crime rate. So a regression co-efficient always does not show what causes what. It just gives us correlation between two variables. 

### Part B. 

The researchers from UPenn came up with a technqiue called intrumental variable in which they use an exogenous shock to police numbers in district one of Washington DC. They came up with a dummy variable of High Alert for days when a high alert was issued due to potential terrorist threat and that led to a higher than usual number of police in the streets of district one. This means that they came up with a variable that only affected the number of police numbers in district one without having any affect on crime rate.

Through the increase in police number due to high alert, the researchers try to look at the casual effect of higher police on crime rate. They compare the crime rate of district one on high alert days with other adjoining districts where police number on streets remain unchanged. So district one had more police on these days as compared to other districts. 

Table 2 gives us coeffcient for high alert days which can be interpreted as a reduction of average 7.3 crimes on high alert days as compared to usual or normal days. This shows that high number of police on high alert days lead to a decrease in crime numbers. 

### Part C 

The main purpose of including this variable is to control for any change in number of public and tourists on high alert days. This approach helps in negating the assumption that crime rate decreases on high alert days because there are fewer people/ tourists on streets and so less oppurtunity for crime. The table shows that even after controlling for metro ridership, we see only a small change in coefficient for high alert day. Thus, high alert is only affecting crime through higher number of police and not through any other means. 

### Part D

From the above regression, we can see that researchers are comparing crime numbers in district one with other districts on high alert days. If only district one witness an increase in police on high alert days, then the decrease in crime rate in district one should be significantly higher than other districts. The table clearly shows this as we look at  (Highalert*District one) coefficient that shows a statistically significant decrease of average 2.8 crimes on high alert days in district one as compared to other districts of Washington that do not show any significant change. The model also includes days fixed effects, day of the week dummies, district fixed effects and district offense fixed effects to isolate the casual effect of police on crime number.  

## Question.3: Clustering and PCA - Wines

In this question we are provided with a dataset of more than six thousand different wines with their 11 chemical properties and two outcome variables quality and color. The task at hand is to employ unsupervised learning techniques and classify these wines into groups/ clusters based on color and quality. 

We start with classfication of wines based on its color. We have two colors; red and white. We plot boxplots of wine color against their chemical properties to see which properties are different across these two colors. We see that amount of chlorides, total sulfur dioxide and volatile acidity are major variables that have different distribution for red and white wines.

```{r, echo = F, cache= TRUE}
wine2= melt(wine[,-c(12,14)], id.vars = "color")
ggplot(data = wine2, aes(x=factor(color), y=value)) + geom_boxplot()+ facet_wrap(~variable, scales = "free") +ggtitle("Component Analysis of Wines on color basis") 
```

### Clustering Kmeans and Kmeanspp

We run a simple kmeans clustering technique and chose centers as 2 because we already know how many groups we have to classify our data into i.e. 2 colors. The result of kmeans clustering in the table below shows that it separates wines based on its color quite successfully. Only 24+68 = 92 wines out of 6,497 wines have been misclassified by this method. The error rate is 1.416 percent.   

```{r fig.align="center", echo = F, cache= TRUE}
X = wine[,-(12:13)]
X = scale(X, center = TRUE, scale = TRUE)
clust1 = kmeans(X, centers = 2, nstart = 25)
pander(table(wine$color, clust1$cluster), caption="Wine Color on Vertical axis vs Cluster group on Horizontal axis for kmeans Clustering")

```

We also plot clusters along with wine colors on a graph with two major differentiating chemical properties. White wines have more total sulfur dioxide content whereas red wines have more chlordies. 

```{r pressure, echo = F, cache= TRUE}
qplot(total.sulfur.dioxide,chlorides, data=wine, color=factor(clust1$cluster))+ggtitle("Cluster Plot with Wine Color and Differentiating features")+ scale_color_manual(breaks = c("1", "2"), values=c("maroon", "grey"))


```

We also employ kmeans++ clustering and it gives us the same result as kmeans clustering which can be seen in the table below.

```{r, echo = F, cache= TRUE}
clust2 = kmeanspp(X, k=2, nstart=50)
pander(table(wine$color, clust2$cluster), caption = "Wine Color on Vertical axis vs Cluster group on Horizontal axis for kmeans++ Clustering")
```

However when using hiearchial clustering, we do not see see clustering based on color of wines. By cutting tree at 2 we get almost all wines sorted into one cluster. Even with cutting tress at 5 or 10 levels, we dont see any balance in the distribution of wines into different clusters. It looks to be a very uneven hiearchial clustering. We try all distance measuring approaches but they all give similar results.

```{r, echo = F}
x_dist = dist(X, method = 'euclidean')
h1 = hclust(x_dist, method ='average')
c1 = cutree(h1, 2)
pander(table(wine$color, c1), caption = "Wine Color on Vertical axis vs Cluster group on Horizontal axis for Hiearchial Clustering")
```


### PCA 
We now use PCA and transform our data accordingly. We get 11 PCAs which are plotted in scree plot below. If we look at first PCA, we see that the main contributers of variation are total sulfur dioxide,chlorides and volatile acidity which are also the main variables that differentiate red from white wines. Thus first PCA is capturing variations between red and white wines. Overall, first 6 PCAs explain 85% of total variation in our dataset of wines.     


```{r, echo = F, cache= TRUE, warning = FALSE}
Y = wine[, -c(12,13,14)]
wine_pc = prcomp(Y, scale.=TRUE)

fviz_screeplot(wine_pc, addlabels = TRUE, ylim = c(0, 35), main = "% of Variation Explained by each PCAs")

fviz_pca_var(wine_pc, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, title = "Variables Contribution in first two PCAs " )

```

We plot our graph for first two PCAs and color the points with color of wines. We can clearly see that PCA1 on the x-axis is differentiating white wines from red wines. So we can say that PCA1 is explaining color variation of wines. 


```{r, echo = F, cache= TRUE}
scores = wine_pc$x

fviz_pca_ind(wine_pc, geom.ind = "point", pointshape = 21, 
             pointsize = 2, 
             fill.ind = wine$color,
             col.ind = "black", 
             palette =  c("red", "grey"), 
             addEllipses = TRUE,
             label = "var",
             col.var = "black",
             repel = TRUE,
             legend.title = "Color") +
  ggtitle("2D PCA-plot for Wine Colors") +
  theme(plot.title = element_text(hjust = 0.5))
```

Now, we use our principal components to run kmeans clustering again to see if we get any better performance than simple kmeans clustering on original scaled data. With clustering on PCAs we see a further reduction in our classification error. Now we only have 37 wines misclassified which equals to an error rate of 0.5%. We can also run a logit model on PCAs but since the question asks us to use only unsupervised learning, we stuck with kmeans clustering.   

```{r, echo = F, cache= TRUE, warning = FALSE}
PCA = scale(scores, center = TRUE, scale = TRUE)
clustpc = kmeans(PCA, centers = 2, nstart = 50)
pander(table(wine$color, clustpc$cluster), caption = "Wine Color on Vertical axis vs Cluster group on Horizontal axis for kmeans Clustering on PCA")
```

From our results we can confidently say that kmeans clustering with PCA gives us a better result in classifying wines based on its color as compared to using clustering on original scaled data. 


### Sorting of Wine based on its Quality

We now try to sort these wines based on their quality. In the data, we have wines with quality ranging from 3 to 9 on a numerical scale. Most of these wines are 5,6 and 7 quality wines.  

```{r, echo = F, cache= TRUE}
hist(wine$quality, xlab = "Wine Quality", main = "Histogram of Wine Qualities")
```

We run the same kmeans clustering but with centers = 7 as now we need to sort wines into 7 clusters/ groups. The confusion matrix shows that there is no pattern of sorting based on quality of wine. For 6 quality wines, the alogrithm places it evenly in almost all clusters. This shows that kmeans is not picking up the differences in wine quality succesfully.


```{r, echo = F, cache= TRUE, warning = FALSE}
clust_ql= kmeans(X, centers=7, nstart=30)
pander(table(wine$quality, clust_ql$cluster), caption = "Wine Quality on Vertical axis vs Cluster group on Horizontal axis for kmeans clustering on original scaled data")
```

We then run kmeans clustering on principal components and sort data into 7 clusters. The consfusion matrix does not look much different than simple kmeans clustering. So even with PCA we are not able to sort wines based on their quality.


```{r, echo = F, cache= TRUE, warning = FALSE}
clustquality = kmeans(PCA, centers= 7, nstart = 70)
pander(table(wine$quality, clustquality$cluster), caption = "Wine Quality on Vertical axis vs Cluster group on Horizontal axis for kmeans Clustering on PCA" )
```

We try to group wines into two groups; high quality and low quality by classifying wines with quality 6 and above as high quality and the rest as low quality wines. We run kmeans clustering again but the results are not encouraging as we see a very high error rate. 

```{r, echo = F, cache= TRUE}
wine1 = mutate(wine, QL = ifelse(quality > 5, "High", "Low"))
pander(table(wine1$QL, clust1$cluster))
```

This clearly shows that the clustering and PCA techniques are not able to sort or identify wines based on its quality. It seems that our approach of euclidean distance measurement is not capturing the differenes in quality of wines. Looking at boxplots of quality against different chemical properties, we do not see much variations for chemical properties across different qualities of wines. Probably the quality of wine is not particularly linked to these 11 chemical properties and hence failure of clustering and PCA to distinguish it. 

```{r, echo = F, cache= TRUE}
wine3= melt(wine[,-c(13,14)], id.vars = "quality")
ggplot(data = wine3, aes(x=factor(quality), y=value)) + geom_boxplot()+ facet_wrap(~variable, scales = "free") +ggtitle("Component Analysis of Wines for different Qualities") 
```

