---
title: "Question 3. Predicting when articles go viral"
output: html_document
---

```{r include=FALSE}
 #### Libraries 
library(tidyverse)
library(mosaic)
library(class)
library(FNN)
library(foreach)
library(class)
library(jtools)
###Data reading
on = read.csv("https://raw.githubusercontent.com/jgscott/ECO395M/master/data/online_news.csv")
###### Adding Viral Binary Variable and Log Shares Variable to the dataset
on$viral = ifelse(on$shares > 1400, 1, 0)
summary(on$viral)
on$lshares =log(on$shares)

```

Description task objectives 


# Approach #1 Working with 
```{r include=FALSE}
### AVG Accuracy model # 0 Null model for comparison (Accuraccy)
n = nrow(on)
n_train = round(0.8*n)  # round to nearest integer
n_test = n - n_train
train_cases = sample.int(n, n_train, replace=FALSE)
test_cases = setdiff(1:n, train_cases)
on_train = on[train_cases,]
null_m=table(on_train$viral)
null_m
ACC_null=null_m[1]/count(on_train)
ACC_null

##### AVG accuracy for Linear Model #1 built manually

lm_shares = do(100)*{
  #	resample data with replacement
    # Split into training and testing sets
  n = nrow(on)
  n_train = round(0.8*n)  # round to nearest integer
  n_test = n - n_train
  train_cases = sample.int(n, n_train, replace=FALSE)
  test_cases = setdiff(1:n, train_cases)
  on_train = on[train_cases,]
  on_test = on[test_cases,]
  y_test   = on$viral[test_cases]
  
    m1 = lm(shares ~ n_tokens_title + n_tokens_content + num_hrefs + 
            num_self_hrefs + num_imgs + num_videos + 
            average_token_length + num_keywords + data_channel_is_lifestyle +
            data_channel_is_entertainment + data_channel_is_bus + 
            data_channel_is_socmed + data_channel_is_tech +
            data_channel_is_world + self_reference_avg_sharess +
            is_weekend + global_rate_negative_words + global_rate_positive_words + title_subjectivity,data=on_train)
  
  pred_test = predict(m1, on_test)
  yhat_test = ifelse((pred_test>1400),yes=1,no=0)
  1-sum(yhat_test != y_test)/n_test 
  
}
ACC_m1=colMeans(lm_shares)

##### AVG accuracy for Linear Model #2 built using KNN with shares
X = dplyr::select(on, n_tokens_title, n_tokens_content, num_hrefs, 
                  num_self_hrefs, num_imgs, num_videos, 
                  average_token_length, num_keywords, self_reference_avg_sharess) 
y = on$shares

k_grid = seq(1, 100, by=5)
err_grid = foreach(k = k_grid,  .combine='c') %do% {
  knn_shares = do(5)*{
    n = length(y)
    n_train = round(0.8*n)
    n_test = n - n_train
    train_ind = sample.int(n, n_train)
    X_train = X[train_ind,]
    X_test = X[-train_ind,]
    y_train = y[train_ind]
    y_test = y[-train_ind]
    y_test_valid=on$viral[-train_ind]
    
    # scale the training set features
    scale_factors = apply(X_train, 2, sd)
    X_train_sc = scale(X_train, scale=scale_factors)
    
    # scale the test set features using the same scale factors
    X_test_sc = scale(X_test, scale=scale_factors)
    
    # Fit two KNN models (notice the odd values of K)
    knn_m2= knn.reg(train=X_train_sc, test= X_test_sc, y=y_train, k=k)
    
   # Calculating classification errors
    phat_test = knn_m2$pred
    yhat_test = ifelse((phat_test>1400),1 , 0)
    sum(yhat_test != y_test_valid)/n_test 
    
  } 
  colMeans(knn_shares)
}
acc_k=1-err_grid
ACC_km2=max(acc_k)
ACC_km2

##### AVG accuracy for Linear Model #3 built using ln of shares 

lm_lshares = do(100)*{
  #	resample data with replacement
  # Split into training and testing sets
  n = nrow(on)
  n_train = round(0.8*n)  # round to nearest integer
  n_test = n - n_train
  train_cases = sample.int(n, n_train, replace=FALSE)
  test_cases = setdiff(1:n, train_cases)
  on_train = on[train_cases,]
  on_test = on[test_cases,]
  y_test   = on$viral[test_cases]
  
  
  m3 = lm(lshares ~ n_tokens_title + n_tokens_content + num_hrefs + 
            num_self_hrefs + num_imgs + num_videos + 
            average_token_length + num_keywords + data_channel_is_lifestyle +
            data_channel_is_entertainment + data_channel_is_bus + 
            data_channel_is_socmed + data_channel_is_tech +
            data_channel_is_world + self_reference_avg_sharess +
            is_weekend + global_rate_negative_words + global_rate_positive_words + title_subjectivity,data=on_train)
  
  pred_test = predict(m3, on_test)
  yhat_test = ifelse(pred_test>log(1400),1 ,0)
  #1-sum(yhat_test != y_test)/n_test ###Accuracy rate
  confusion_out = table(y = on_test$viral, yhat = yhat_test)
}
CM=colMeans(lm_lshares)
tp=CM[4]
tn=CM[1]
fp=CM[3]
fn=CM[2]
tpr = tp / (tp + fn)
tpr
fpr = fp / (fp + tn)
fpr
ACC_m3 = (tp+tn)/(tp+tn+fp+fn)
ACC_m3
ERR_m3=1-ACC_m3
ERR_m3


```



Present a table with ACC of the models and comment, then present the Conf matrix for the best model and the requested ratios. Comment. 


lshares ~ n_tokens_title + n_tokens_content + num_hrefs + 
            num_self_hrefs + num_imgs + num_videos + 
            average_token_length + num_keywords + data_channel_is_lifestyle +
            data_channel_is_entertainment + data_channel_is_bus + 
            data_channel_is_socmed + data_channel_is_tech +
            data_channel_is_world + self_reference_avg_sharess +
            is_weekend + global_rate_negative_words + global_rate_positive_words +       title_subjectivity




